---
title: "Quantified Self-movement Data Analysis"
output: 
  html_document:
    keep_md: true
---

#### *Fernaldo R. Winnerdy*
#### *28-Sep-2018*

The project is about predicting the forms of exercise based on the measurements of accelerometer on 6 different subjects. The test subjects were to performed belt, forearm, arm, and dumbell exrecises in a correct form and four different commonly used wrong forms. Base on the data, we try to predict the forms that a person do.

### Data import
First, we import the data to R (data were downloaded beforehand) and load the required library.

```{r import}
library(caret)
library(rpart)
setwd('/home/fernaldo/Projects/Others/Programming/R-specialization/08-Practical_machine_learning/Week-04')
training <- read.csv('pml-training.csv')
testing <- read.csv('pml-testing.csv')
```

### Cleaning data

Here, we clean the data (both training and testing) such that it includes only the variables from the accelerometer, as well as removing the variables with NA.

```{r cleaning}
accel_index <- grep(pattern = 'accel', names(training))
set_training <- training[, c(accel_index, 160)]
set_training <- set_training[, colSums(is.na(set_training)) == 0]
set_testing <- testing[, c(accel_index, 160)]
set_testing <- set_testing[, colSums(is.na(set_testing)) == 0]
```

We end up with only 17 variables in both training and testing set

### Cross validation separation

Since we need out of sample error estimate using cross-validation, we need to further split the training set into 2, 70% into a training set and 30% into a validation set.

```{r split}
inTrain <- createDataPartition(y = set_training$classe, p=0.7, list=FALSE)
tra <- set_training[inTrain, ]
val <- set_training[-inTrain, ]
```

For cross validation, we separate the data into k=5 folds.

```{r cross}
Control <- trainControl(method = "cv", number = 5)
```

### Model fitting

Random forest model is chosen. Accuracy of this model is computed based on the k-fold cross validation method.

```{r model, cache=TRUE}
fit <- train(classe ~ ., method="rf", data = tra, trControl = Control)
fit$resample
confusionMatrix.train(fit)
```

The average accuracy is ~94% based on the validation technique that we used.  

### Cross Validation accuracy

The out of sample accuracy (using validation data):

```{r validate}
confusionMatrix(predict(fit, val), val$classe)
```

Given 98.5% accuracy of the prediction using the validation set, the expected out of sample error is ~2.5%.

### Test result
The test prediction results are as follows

```{r result}
result <- predict(fit, set_testing)
result
```

The above results were used to predict the classe variables of the testing set in the following project quiz. 